A Generalized Sphere Theorem for Positively Curved Combinatorial 3-Manifolds

### 2013-06-25 paper/paper.tex
messed with tables
journal specific stuff added

\documentclass[12pt]{article}

\usepackage[margin=1in]{geometry}  % set the margins to 1in on all sides
\usepackage{deluxetable}             
\usepackage{graphicx}              % to include figures
\usepackage{amsmath}               % great math stuff
\usepackage{amsfonts}              % for blackboard bold, etc
\usepackage{amsthm}                % better theorem environments
\usepackage{url}

% various theorems, numbered by section

\newtheorem{thm}{Theorem}[section]
\newtheorem{lem}[thm]{Lemma}
\newtheorem{prop}[thm]{Proposition}
\newtheorem{cor}[thm]{Corollary}
\newtheorem{conj}[thm]{Conjecture}
\newtheorem{dfn}[thm]{Definition}

\DeclareMathOperator{\id}{id}

\newcommand{\bd}[1]{\mathbf{#1}}  % for bolding symbols
\newcommand{\RR}{\mathbb{R}}      % for Real numbers
\newcommand{\ZZ}{\mathbb{Z}}      % for Integers
\newcommand{\col}[1]{\left[\begin{matrix} #1 \end{matrix} \right]}
\newcommand{\comb}[2]{\binom{#1^2 + #2^2}{#1+#2}}

@@

\begin{thm}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and diameter greater than $\frac{\pi}{2\sqrt{\delta}}$. Then, $M$ is homeomorphic to a sphere.
\label{thm:grove_shiohama}
\end{thm}

@@

\begin{thm}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with sectional curvature $K \geq \delta > 0$. Then the diameter of $M$ is at most $\frac{\pi}{\sqrt{\delta}}$.
\label{thm:bonnet_myers}
\end{thm}

@@

\begin{dfn}The \textbf{edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length (as a PL-path in the standard PL-metric) of an edge-path from $v$ to $w$. We denote this quantity $d_1(v,w)$. The \textbf{edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$. 
\end{dfn}

@@

\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
\label{thm:discrete_BM}
\end{thm}

@@

\begin{dfn}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ are vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$. The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a \textbf{hop} from $v_1$ to $v_2$. See Figure \ref{fig:hop} for an illustration.
\end{dfn}

@@

\begin{dfn}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$. We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a \textbf{jump} from $v_1$ to $v_2$. For an illustration, see Figure \ref{fig:jump}.
\end{dfn}

@@

\begin{lem}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
\end{lem}

@@

\begin{thm}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
\label{thm:discrete_BM_expanded_paths}
\end{thm}

@@

\begin{thm} Let $M$ be a connected, positively curved, combinatorial 3-manifold. If $M$ has $diam(M)>J$ then it is homeomorphic to a 3-sphere. Moreover, this bound is sharp and equal to half the maximum diameter which occurs for any such $M$.
\label{thm:discrete_GS}
\end{thm}

@@

\begin{thm} If $M$ is a connected, positively curved, combinatorial 3-manifold satisfying the bound $diam_1(M) > 3$ then $M$ is homeomorphic to the 3-sphere. Moreover, this bound is sharp.
\end{thm}

@@

\begin{thm}(Unique Extension) If two minimal paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of a positively curved 3-manifold $M$ and $P_1$ and $P_2$ have the same length, then $P_1 = P_2$. This result is false if $\mathcal{P}$ is replaced by $\mathcal{P}_1$.
\end{thm}

@@

\input{bigtable.tex}

>>>>

% EJC papers *must* begin with the following two lines. 
\documentclass[12pt]{article}
\usepackage{e-jc}

% Please remove all other commands that change parameters such as
% margins or pagesizes.

% only use standard LaTeX packages
% only include packages that you actually need

% we recommend these ams packages
\usepackage{amsthm,amsmath,amssymb}

% we recommend the graphicx package for importing figures
\usepackage{graphicx}

% use this command to create hyperlinks (optional and recommended)
\usepackage[colorlinks=true,citecolor=black,linkcolor=black,urlcolor=blue]{hyperref}

% use these commands for typesetting doi and arXiv references in the bibliography
\newcommand{\doi}[1]{\href{http://dx.doi.org/#1}{\texttt{doi:#1}}}
\newcommand{\arxiv}[1]{\href{http://arxiv.org/abs/#1}{\texttt{arXiv:#1}}}

% all overfull boxes must be fixed; 
% i.e. there must be no text protruding into the margins

% declare theorem-like environments
\theoremstyle{plain}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{fact}[theorem]{Fact}
\newtheorem{observation}[theorem]{Observation}
\newtheorem{claim}[theorem]{Claim}

\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{open}[theorem]{Open Problem}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{question}[theorem]{Question}

\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{note}[theorem]{Note}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\begin{theorem}[Grove-Shiohama] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with section curvature $K \geq \delta > 0$ and diameter greater than $\frac{\pi}{2\sqrt{\delta}}$. Then, $M$ is homeomorphic to a sphere.
\label{thm:grove_shiohama}
\end{theorem}

@@

\begin{theorem}[Bonnet-Myers] Let $M$ be a complete, connected, $n$-dimensional Riemannian manifold with sectional curvature $K \geq \delta > 0$. Then the diameter of $M$ is at most $\frac{\pi}{\sqrt{\delta}}$.
\label{thm:bonnet_myers}
\end{theorem}

@@

\begin{definition}The \textbf{edge-distance} between two vertices $v$ and $w$ in an abstract simplicial complex $M$ is the minimum length (as a PL-path in the standard PL-metric) of an edge-path from $v$ to $w$. We denote this quantity $d_1(v,w)$. The \textbf{edge-diameter} of $M$, written as $diam_1(M)$, is the maximum of $d_1(v,w)$ over all pairs of vertices $v$ and $w$ in $M$. 
\end{definition}

@@

\begin{theorem}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam_1(M)\leq 5$. Moreover, this bound is sharp.
\label{thm:discrete_BM}
\end{theorem}

@@

\begin{definition}[Hops] Suppose $\tau$ is a 2-simplex in $M$ and $v_1$ and $v_2$ are vertices in $M$ such that $v_1*\tau$ and $v_2*\tau$ are 3-simplices in $M$. The PL-path from $v_1$ through the barycenter of $\tau$ and ending on $v_2$ will be called a \textbf{hop} from $v_1$ to $v_2$. See Figure \ref{fig:hop} for an illustration.
\end{definition}

@@

\begin{definition}[Jumps] Suppose there are edges $e_1$ and $e_2$ and vertices $v_1$ and $v_2$ in $M$ so that $e_1*e_2$ is a 3-simplex in $M$ and $v_1*e_1$ and $v_2*e_2$ are 2-simplices in $M$. We call the PL-path from $v_1$ through the barycenters of $e_1$ and $e_2$ and ending on $v_2$ a \textbf{jump} from $v_1$ to $v_2$. For an illustration, see Figure \ref{fig:jump}.
\end{definition}

@@

\begin{lemma}[Trout] If $v$ and $w$ are vertices in a connected positively curved combinatorial 3-manifold $M$ then $d(v,w) \in \{0, 1, H, 2, J, 3, 2H, 4, 2J \}$.
\end{lemma}

@@

\begin{theorem}[Trout] For any connected positively curved combinatorial 3-manifold $M$ we have $diam(M) \leq 2J$.
\label{thm:discrete_BM_expanded_paths}
\end{theorem}

@@

\begin{theorem} Let $M$ be a connected, positively curved, combinatorial 3-manifold. If $M$ has $diam(M)>J$ then it is homeomorphic to a 3-sphere. Moreover, this bound is sharp and equal to half the maximum diameter which occurs for any such $M$.
\label{thm:discrete_GS}
\end{theorem}

@@

\begin{theorem} If $M$ is a connected, positively curved, combinatorial 3-manifold satisfying the bound $diam_1(M) > 3$ then $M$ is homeomorphic to the 3-sphere. Moreover, this bound is sharp.
\end{theorem}

@@

\begin{theorem}(Unique Extension) If two minimal paths $P_1$ and $P_2$ in $\mathcal{P}$ pass through the same first two simplices of a positively curved 3-manifold $M$ and $P_1$ and $P_2$ have the same length, then $P_1 = P_2$. This result is false if $\mathcal{P}$ is replaced by $\mathcal{P}_1$.
\end{theorem}

@@

% \input{bigtable.tex}
